### 2025-11-11 sections/2-methods.tex
made minor edits and fixed typos

        \If{$|C| > 1$ \textbf{and} $\mathcal{C}$ satisfies $criteria$}
        \State $c \Leftarrow$ geometric median of $\mathcal{C}$
        \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
        \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
        \State $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
        \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
        \State \Call{Partition}{L, criteria}
        \State \Call{Partition}{R, criteria}
        \EndIf
    \end{algorithmic}
\end{algorithm}

@@

We begin by using \textsc{TreeSearch} on the root cluster $\mathcal{R}$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|\textbf{X}|$.
If the returned set $\mathcal{Q}$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $\mathcal{Q}$ is non-empty.
Now, so long as $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}|}} \big)^\frac{1}{\mu}$
where $\mu$ is the harmonic mean of the LFD of the clusters in $\mathcal{Q}$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
Once $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| \geq k$, we perform \textsc{LeafSearch} to find the $k$ nearest neighbors.
\begin{algorithm}[h]
    \caption{\textsc{TreeSearch}($\mathcal{C}$, $q$, $r$)}
    \label{alg:methods:rnn-search:tree-search}
    \begin{algorithmic}
        \Require $\mathcal{C}$, a cluster
        \Require $q$, a query
        \Require $r$, a search radius
        \If{$\delta^+_C \leq r$}
            \State \textbf{return} $\{C\}$
        \EndIf
        \State $[L, R]$ $\Leftarrow$ \textit{children} of $\mathcal{C}$
        \State \textbf{return} \Call{TreeSearch}{$L, q, r$} $\cup$ \Call{TreeSearch}{$R, q, r$}
    \end{algorithmic}
\end{algorithm}
\vspace{-0.2in}

\begin{algorithm}[h]
    \caption{\textsc{LeafSearch}($\mathcal{Q}$, $q$, $r$)}
    \label{alg:methods:rnn-search:leaf-search}
    \begin{algorithmic}
        \Require $f$, a distance function
        \Require $\mathcal{Q}$, a set of clusters
        \Require $q$, a query
        \Require $r$, a search radius
        \State $H \Leftarrow \emptyset$
        \For{$C \in Q$}
            \For{$p \in C$ \textbf{if} $f(p, q) \leq r$}
                \State $\mathcal{H}$ $\Leftarrow$ $H \cup \{p\}$
            \EndFor
        \EndFor
        \State \textbf{return} $\mathcal{H}$
    \end{algorithmic}
\end{algorithm}
\vspace{-0.2in}

@@

We begin by letting $\mathcal{Q}$ be a list of 3-tuples $(\mathcal{C}, \delta^{+}_{\mathcal{C}}, m)$, where $m$ is the multiplicity of $\mathcal{C}$ in $\mathcal{Q}$.
$\mathcal{Q}$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $\mathcal{H}$ starts with the center of $\mathcal{R}$, where $\mathcal{R}$ is the root of $\mathcal{T}$.
We then use QuickSelect, modified to account for multiplicities and to reorder $\mathcal{Q}$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $\mathcal{Q}$ such that $\left| B(q, \tau) \right| \geq k$ (including the points in $\mathcal{H}$).
Since this step may require a binary search of $\mathcal{Q}$ and reordering $\mathcal{Q}$ takes linear time, this version of QuickSelect has $\mathcal{O}\left(|\mathcal{Q}| \log |\mathcal{Q}|\right)$ time complexity.

>>>>

        \If{$|\mathcal{C}| > 1$ \textbf{and} $\mathcal{C}$ satisfies $criteria$}
        \State $c \Leftarrow$ geometric median of $\mathcal{C}$
        \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in \mathcal{C}$
        \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in \mathcal{C}$
        \State $\mathcal{L} \Leftarrow \{x \ | \ x \in \mathcal{C} \land f(l, x) \le f(r, x)\}$
        \State $\mathcal{R} \Leftarrow \{x \ | \ x \in \mathcal{C} \land f(r, x) < f(l, x)\}$
        \State \Call{Partition}{$\mathcal{L}$, criteria}
        \State \Call{Partition}{$\mathcal{R}$, criteria}
        \EndIf
    \end{algorithmic}
\end{algorithm}

@@

We begin by using \textsc{TreeSearch} (Alg~\ref{alg:methods:rnn-search:tree-search}) on the root cluster $\mathcal{R}$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|\textbf{X}|$.
If the returned set $\mathcal{Q}$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $\mathcal{Q}$ is non-empty.
Now, so long as $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}|}} \big)^\frac{1}{\mu}$
where $\mu$ is the harmonic mean of the LFD of the clusters in $\mathcal{Q}$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
Once $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| \geq k$, we perform \textsc{LeafSearch} (Alg~\ref{alg:methods:rnn-search:leaf-search}) to find the $k$ nearest neighbors.
\begin{algorithm}[h]
    \caption{\textsc{TreeSearch}($\mathcal{C}$, $q$, $r$)}
    \label{alg:methods:rnn-search:tree-search}
    \begin{algorithmic}
        \Require $\mathcal{C}$, a cluster
        \Require $q$, a query
        \Require $r$, a search radius
        \If{$\delta^+_\mathcal{C} \leq r$}
            \State \textbf{return} $\{\mathcal{C}\}$
        \EndIf
        \State $[\mathcal{L}, \mathcal{R}]$ $\Leftarrow$ \textit{children} of $\mathcal{C}$
        \State \textbf{return} \Call{TreeSearch}{$\mathcal{L}, q, r$} $\cup$ \Call{TreeSearch}{$\mathcal{R}, q, r$}
    \end{algorithmic}
\end{algorithm}
\vspace{-0.2in}

\begin{algorithm}[h]
    \caption{\textsc{LeafSearch}($\mathcal{Q}$, $q$, $r$)}
    \label{alg:methods:rnn-search:leaf-search}
    \begin{algorithmic}
        \Require $f$, a distance function
        \Require $\mathcal{Q}$, a set of clusters
        \Require $q$, a query
        \Require $r$, a search radius
        \State $\mathcal{H} \Leftarrow \emptyset$
        \For{$\mathcal{C} \in \mathcal{Q}$}
            \For{$p \in \mathcal{C}$ \textbf{if} $f(p, q) \leq r$}
                \State $\mathcal{H}$ $\Leftarrow$ $\mathcal{H} \cup \{p\}$
            \EndFor
        \EndFor
        \State \textbf{return} $\mathcal{H}$
    \end{algorithmic}
\end{algorithm}
\vspace{-0.2in}

@@

We begin by letting $\mathcal{Q}$ be a list of 3-tuples $(\mathcal{C}, \delta^{+}_{\mathcal{C}}, m)$, where $m$ is the multiplicity of $\mathcal{C}$ in $\mathcal{Q}$.
$\mathcal{Q}$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $\mathcal{H}$ starts with the center of $\mathcal{R}$.
We then use QuickSelect, modified to account for multiplicities and to reorder $\mathcal{Q}$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $\mathcal{Q}$ such that $\left| B(q, \tau) \right| \geq k$ (including the points in $\mathcal{H}$).

### 2025-11-11 sections/3-datasets.tex
made minor edits and fixed typos

HNSW, ANNOY, and FAISS have various hyper-parameters that can be tuned to trade off throughput, recall, and indexing time.
We allow for a hyper-parameter search (in the ranges recommended by the documentation accompanying the respective libraries) to find the configuration with the highest throughput that still achieves perfect recall (or as close to it as possible).
To make the comparison fair, we also allow CAKES to select the fastest of its algorithms by testing each with a small number of the centers of clusters at low depth as representative queries.

>>>>

HNSW, ANNOY, and FAISS have various hyper-parameters that can be tuned to trade off throughput, recall, and indexing time.
We allow for a hyper-parameter search (in the ranges recommended by the documentation accompanying the respective libraries) to find the configuration with the highest throughput that still achieves perfect recall (or as close to it as possible).
To make the comparison fair, we also allow CAKES to select the fastest of its algorithms by testing each with a small number of the centers of clusters at low depth in the tree as representative queries.

### 2025-11-11 sections/4-results.tex
made minor edits and fixed typos

On Silva (Fig.~\ref{fig:results:silva-scaling}), all CAKES algorithms initially exhibit throughput that decreases approximately linearly with cardinality, but then level off at higher cardinalities.
DFS is consistently the fastest algorithm on this dataset.

@@

  \vskip 0.1in
  \includegraphics[width=0.75\columnwidth]{plots/legend.png}

>>>>

On Silva (Fig.~\ref{fig:results:silva-scaling}), all CAKES algorithms initially exhibit throughput that decreases approximately linearly with cardinality, but then levels off at higher cardinalities.
DFS is consistently the fastest algorithm on this dataset.

@@

  \vskip 0.1in
  \includegraphics[width=0.75\columnwidth]{plots/legend_only_cakes.png}
